Safe Memory Regions for Big Data Processing

### 2016-06-01 paper/type-inference.tex
type inference

\name's region type system imposes a heavy annotation burden, and the
cost of manually annotating C\# standard libraries with region types
is prohibitive. We now present our region type inference algorithm
that completely eliminates the need to write region type annotations,
except on some higher-order functions. Formally, the type inference
algorithm is an elaboration function from programs in $\absof{\FB}$
(i.e., \FB without region types, but with \C{letregion} and \C{open}
expressions, similar to the language introduced in
\S~\ref{sec:overview}) to programs in \FB.

\textbf{TODO: Mention that constraints are in the theory of partial orders?}

@@

>>>>

\textbf{TO DO:}
\begin{itemize}
\item Add destructive object-field-update \emph{x.f := y} to the language.
\item Add if-then-else to the language.
\item Note that the claim that all region variables will be unified with some concrete
region fails in the case of a method that recursively calls itself (and, hence, does not
terminate). In this case, the region variables in the return type are not constrained.
Similarly for non-terminating mutual recursion. These are uninteresting (in essence, the
return type can be taken to be unit instead).
\item Discuss the safety of virtual method invocation, which provides a limited form
of existentially quantifying regions and instantiating (unpacking) them because a
derived class may have more region parameters than a base class.
\item Discuss the subtyping rule used to check an overriding method (in the static
semantics).
\item Mention that constraints are in the theory of partial orders?
\end{itemize}

\name's region type system imposes a heavy annotation burden, and
manually annotating C\# standard libraries with region types
can be tedious. We now present our region type inference algorithm
that eliminates the need to write region type annotations.
% except on some higher-order functions.
Formally, the type inference
algorithm is an elaboration function from programs in $\absof{\FB}$
(i.e., \FB without region types, but with \C{letregion} and \C{open}
expressions, similar to the language introduced in
\S~\ref{sec:overview}) to programs in \FB.

@@

The first step involves the following two mutually dependent ideas.

\emph{Region Parametrization}. For every class \C{C}, we identify
a sequence of region parameters $\pi_0, \cdots \pi_n$ that \C{C}
should be parametric over.

\emph{Instantiation of Region Parametric Classes}. For every region
parametric class $\C{C}\langle \pi_0, \cdots, \pi_n \rangle$ so determined,
we replace every instance of class \C{C} in the program by an instance
$\C{C}\langle \rho_0, \cdots, \rho_n \rangle$, where $\rho_0, \cdots, \rho_n$
are fresh identifiers.

In particular, we determine the sequence of region parameters for a class \C{C}
by starting with the sequence of region parameters required for the type of
the base class of \C{C}, and adding fresh region parameters for the type of every field
in the class.
The class \C{Object} has a single region parameter $\pi_0$ (the allocation region).
